---
title: "Code for Manuscript: Salt marsh runnels alter fish activity but not benthic macroinvertebrate abundance and diversity"
authors: "Amanda Wik and Nathalie Sommer"
date: "March 2026"
---

#Prepping R packages

```{r load packages}

library(vegan)
library(ape)
library(dplyr)
library(tidyverse)
library(textshape)
library(knitr)
library(ggplot2)
library(lme4)
library(ggeffects)  
library(lmerTest)
library(ggfortify)
library(MuMIn)
library(pbapply)
library(AICcmodavg)
library(rsq)
library(gtable)
library(gt)
library(TMB)
library(glmmTMB)
library(DHARMa)
library(ggrepel)
library(car)
library(lme4)
library(betareg)

```

# Data cleaning
## Inverts

```{r load invertebrate data}


combined_invert_data <- read.csv('Invertebrate_Data_2024.csv', 
                      header=TRUE, stringsAsFactors=FALSE)

# Convert character variables to factors
combined_invert_data$Treatment <- factor(combined_invert_data$Treatment)
combined_invert_data$Water_Level <- factor(combined_invert_data$Water_Level)

#Adding Site_ID as rownames
combined_invert_data <- combined_invert_data %>%
  mutate(Site_ID = paste(Site, ID, sep = "")) %>% 
  column_to_rownames("Site_ID")

```

```{r calculating Infauna and Epifauna abundance, richness and diversity}

#Calculating Abundance (total number of invertebrates)
combined_invert_data <- combined_invert_data %>%
  mutate(Total_Epifauna = rowSums(across(Amphipods:Stratiomydae), na.rm = FALSE))
combined_invert_data <- combined_invert_data %>%
  mutate(Total_Infauna = rowSums(across(Capitella:Unknown_Infauna), na.rm = FALSE))

#Calculating Richness
combined_invert_data$Epifauna_Richness <- as.integer(specnumber(select(combined_invert_data, Amphipods:Stratiomydae)))
combined_invert_data$Infauna_Richness <- as.integer(specnumber(select(combined_invert_data, Capitella:Unknown_Infauna)))

#Calculating Shannon diversity
combined_invert_data$Epifauna_Shannon <- diversity(select(combined_invert_data, Amphipods:Stratiomydae), index = "shannon")
combined_invert_data$Infauna_Shannon <- diversity(select(combined_invert_data, Capitella:Unknown_Infauna), index = "shannon")

```

## Fish

``` {r loading fish data and creating field summary data}

#Loading in the fish data

fish_diet_data <- read.csv('Fish_Gut_Data_2025.csv', 
                      header=TRUE, stringsAsFactors=FALSE)


fish_survey_data <- read.csv('Fish_Survey_Data_2024.csv', 
                      header=TRUE, stringsAsFactors=FALSE)

fish_size_data <- read.csv('Fish_Size_Data_2024.csv', 
                      header=TRUE, stringsAsFactors=FALSE)

# Convert character variables to factors
fish_survey_data$Tide <- factor(fish_survey_data$Tide)
fish_survey_data$Trap_Location <- factor(fish_survey_data$Trap_Location)
fish_survey_data$Treatment <- factor(fish_survey_data$Treatment)
fish_survey_data$Trap_Type <- factor(fish_survey_data$Trap_Type)

#Combining lab and field fish data
combined_fish_data <- merge(
  fish_survey_data,
  fish_diet_data,
  by = c("Site", "ID", "Tide"),
  all.y = TRUE
)

#Creating a dataframe of mean weight, mean length, and number of fish by trap (plot)
fish_stats_by_plot <- fish_size_data %>%
  group_by(Site, ID, Tide) %>%
  mutate(
    Mean_Weight_by_Trap = mean(Fish_Weight, na.rm = TRUE),
    Mean_Length_by_Trap = mean(Fish_Length, na.rm = TRUE),
    Total_Number_Fish_by_Trap = n()
  ) %>%
  ungroup() %>%
  select(-Fish_ID:-Fish_Weight) %>%
  distinct() 

#Creating a dataframe of mean weight, length, and number of fish by site
fish_stats_by_site <- fish_size_data %>%
  group_by(Site, Tide) %>%
  mutate(
    Mean_Weight_by_Trap = mean(Fish_Weight, na.rm = TRUE),
    Mean_Length_by_Trap = mean(Fish_Length, na.rm = TRUE),
    Total_Number_Fish_by_Trap = n()
  ) %>%
  ungroup() %>%
  select(-ID, -Fish_ID:-Fish_Weight) %>%
  distinct()

#Creating count data for each plot that includes zeroes where no fish were caught
fish_field_data <- fish_survey_data %>%
  left_join(fish_stats_by_plot %>% select(Site, ID, Tide, Total_Number_Fish_by_Trap),
            by = c("Site", "ID", "Tide")) %>%
  mutate(Total_Number_Fish_by_Trap = replace_na(Total_Number_Fish_by_Trap, 0))

#Adding site data to fish field size dataframe
fish_field_size_data <- fish_size_data %>%
  left_join(fish_survey_data %>% select(Site:Treatment),
            by = c("Site", "ID", "Tide")) 

```

```{r Prepping fish diet data, adding richness, diversity, and other columns}

#Adding total crab column
combined_fish_data <- combined_fish_data %>%
  mutate(Total_Crab = Crabs_Whole + Crab_Parts)

# Calculate species richness
combined_fish_data$Diet_Richness <- specnumber(select(combined_fish_data, c(Algae:Beetle_Larvae, Diptera:Total_Crab)))

# Calculate Shannon diversity
combined_fish_data$Diet_Diversity_Shannon <- combined_fish_data %>%
  select(Algae:Beetle_Larvae, Diptera:Total_Crab) %>%
  mutate(across(everything(), as.numeric)) %>%
  as.matrix() %>%
  diversity(index = "shannon")

#Adding Fish IDs as row names
combined_fish_data <- combined_fish_data %>%
  mutate(Site_ID_Tide_Fish = paste(Site, ID, Tide, Fish_ID, sep = "")) %>% 
  column_to_rownames("Site_ID_Tide_Fish")

```

```{r data preparation - excluding runnel traps}

# Some traps were placed inside runnels for the restored sites, so creating a data frame that only includes traps placed on the marsh platform

fish_field_data_no_traps_in_runnels <- fish_field_data %>% filter(Trap_Location != "Runnel")
fish_field_size_data_no_traps_in_runnels <- fish_field_size_data %>% filter(Trap_Location != "Runnel")

# Checking the distribution of traps

table(fish_field_data_no_traps_in_runnels$Trap_Type, fish_field_data_no_traps_in_runnels$Treatment)
table(fish_field_data_no_traps_in_runnels$Trap_Type, fish_field_data_no_traps_in_runnels$Trap_Location)

# Summary
cat("Data structure for main analyses (excluding traps in runnels):\n")
cat("Total observations:", nrow(fish_field_data_no_traps_in_runnels), "\n")
cat("Runnel sites:", sum(fish_field_data_no_traps_in_runnels$Treatment == "Runnel"), "\n")
cat("Non-runnel sites:", sum(fish_field_data_no_traps_in_runnels$Treatment == "Control"), "\n")

```

#Epifauna NMDS

```{r Epi: prepping the data for NMDS}

# Separating environmental data and species data; first is creating a species dataframe
epifauna_spp <- combined_invert_data %>% select(Amphipods:Stratiomydae)
# Removing row LBSC6 because it contains only zeroes
epifauna_spp <- epifauna_spp[!rownames(epifauna_spp) %in% "LBSC6", ]

#Creating an environmental dataframe
epifauna_treatment <- combined_invert_data %>% 
  select("Site", "ID", "Treatment") %>% 
  mutate(Site_ID = paste(Site, ID, sep = "")) %>% 
  select(-Site, -ID)
  
# Removing row LBSC6 because it contains only zeroes
epifauna_treatment <- epifauna_treatment[!rownames(epifauna_treatment) %in% "LBSC6", ]

```

```{r Epi: conducting the NMDS}
# Performing nmds with environmental (treatment) and species vectors
epifauna_mds <- metaMDS(epifauna_spp, distance = "bray", autotransform = FALSE)

epifauna_sppfit <- envfit(epifauna_mds, epifauna_spp, permutations = 999) # Fit species vectors

epifauna_mds  # Stress value is less than 0.2, which is good.

```

```{r Epi: saving the NMDS results}

# Saving the results from the nmds and grouping the data by environmental variables
epifauna_site_scrs <- as.data.frame(scores(epifauna_mds, display = "sites"))  # Saving NMDS results into data frame

epifauna_site_scrs <- cbind(epifauna_site_scrs, Treatment = epifauna_treatment$Treatment) # Making treatment a grouping variable and saving to data frame

```

```{r Epi: creating a dataframe with the signifcant species contributions}

# Saving species data from nmds to dataframe
epifauna_spp_scrs <- as.data.frame(scores(epifauna_sppfit, display = "vectors"))  # Saving species values into dataframe

epifauna_spp_scrs <- cbind(epifauna_spp_scrs, Species = rownames(epifauna_spp_scrs))  # Adding species names to dataframe

epifauna_spp_scrs <- cbind(epifauna_spp_scrs, pval = epifauna_sppfit$vectors$pvals) # Adding p values to dataframe to later select species which are significant

sig_epifauna_spp_scrs <- subset(epifauna_spp_scrs, pval<=0.05) # Showing only significant species (using 0.05 as cutoff)

head(epifauna_spp_scrs)  # Viewing data frame

```

```{r Epi: PERMANOVA}

# PERMANOVA (adonis2) 
permanova_result_epifauna <- adonis2(epifauna_spp ~ Treatment, data = epifauna_treatment, permutations = 999)
print(permanova_result_epifauna)

```

```{r Epi: SIMPER}

# SIMPER looking for species contributing most to difference between treatment groups
mat_bar_epifauna_spp <- as.matrix(epifauna_spp)
simp_treatment_epifauna <- simper(mat_bar_epifauna_spp, epifauna_treatment$Treatment)
simp_treatment_epifauna
summary(simp_treatment_epifauna, ordered = TRUE, digits = 3)

```

#Infauna NMDS

```{r In: prepping the data for NMDS}

# Separating environmental data and species data; first is the species dataframe
infauna_spp <- combined_invert_data %>% select(Capitella:Unknown_Infauna) %>% na.omit() 
rows_to_remove <- c("LBSL3", "LBNE4", "LBNG6", "LBSC1", "LBSC2", "LBSC3")  # Adding all the row names to be removed here because of zeroes
infauna_spp <- infauna_spp[!rownames(infauna_spp) %in% rows_to_remove, ] # Removing those rows

# Now making the treatment dataframe
infauna_treatment <- combined_invert_data %>% 
  drop_na(Total_Infauna) %>% 
  select("Site","ID","Treatment") %>% 
  mutate(Site_ID = paste(Site, ID, sep = "")) %>% 
  select(-Site, -ID) 

# Removing rows with all zeroes
infauna_treatment <- infauna_treatment[!rownames(infauna_treatment) %in% rows_to_remove, ]

```

```{r In: conducting the NMDS}
# Performing nmds with environmental (treatment) and species vectors
infauna_mds <- metaMDS(infauna_spp, distance = "bray", autotransform = FALSE)

infauna_sppfit <- envfit(infauna_mds, infauna_spp, permutations = 999) # Fitting species vectors

infauna_mds  # Stress value is less than 0.2 (0.1293137), which is good. Shows how easy it was to condense multidimensional data into two dimensional space

```

```{r In: saving the NMDS results}

# Saving the results from the nmds and grouping the data by environmental variables
infauna_site_scrs <- as.data.frame(scores(infauna_mds, display = "sites"))  # Saving NMDS results into dataframe
infauna_site_scrs <- cbind(infauna_site_scrs, Treatment = infauna_treatment$Treatment)  # Making treatment a grouping variable and saving to dataframe

head(infauna_site_scrs)  # Viewing the dataframe

```

```{r In: creating a data frame with the signifcant species contributions}

# Saving species data from nmds to dataframe
infauna_spp_scrs <- as.data.frame(scores(infauna_sppfit, display = "vectors"))  # Saving species values into dataframe
infauna_spp_scrs <- cbind(infauna_spp_scrs, Species = rownames(infauna_spp_scrs))  # Adding species names to dataframe
infauna_spp_scrs <- cbind(infauna_spp_scrs, pval = infauna_sppfit$vectors$pvals) # Adding p values to dataframe to select species which are significant
sig_infauna_spp_scrs <- subset(infauna_spp_scrs, pval<=0.05) # Showing only significant species (using 0.05 as cutoff)
head(infauna_spp_scrs)  # Viewing dataframe

```


```{r In: PERMANOVA}

# PERMANOVA (adonis2) 
permanova_result_infauna <- adonis2(infauna_spp ~ Treatment, data = infauna_treatment, permutations = 999)
print(permanova_result_infauna)

```

```{r In: SIMPER}

# SIMPER looking for species contributing most to difference between treatment groups
mat_bar_infauna_spp <- as.matrix(infauna_spp)
simp_treatment_infauna <- simper(mat_bar_infauna_spp, infauna_treatment$Treatment)
simp_treatment_infauna
summary(simp_treatment_infauna, ordered = TRUE, digits = 3)

```

#Invertebrate Abundance
```{r glmmTMB modelling (nbinomial) for total epifauna}

#epifauna model

epi_number_glmmtmb <- glmmTMB(Total_Epifauna ~ Treatment + Water_Level + Veg_Cover + (1 | Site),
                               data = combined_invert_data,
                               family = nbinom2(link = "log"))

#Checking the residuals to see if the q-q plot looks okay, and to see if there is any heteroscedasticity
simulateResiduals(epi_number_glmmtmb, plot=TRUE)

#Checking the results and estimates
summary(epi_number_glmmtmb)
confint(epi_number_glmmtmb)
r.squaredGLMM(epi_number_glmmtmb)

```

```{r glmmTMB modelling (nbinomial) for total infauna}

#infauna model

in_number_glmmtmb <- glmmTMB(Total_Infauna ~ Treatment + Water_Level + Veg_Cover + (1 | Site),
                               data = combined_invert_data,
                               family = nbinom2(link = "log"))

#Checking the residuals to see if the q-q plot looks okay,and if there is any heteroscedasticity
simulateResiduals(in_number_glmmtmb, plot=TRUE)

#Checking the results and estimates
summary(in_number_glmmtmb)
confint(in_number_glmmtmb)
r.squaredGLMM(in_number_glmmtmb)

```

#Invertebrate Richness
```{r glmmTMB modelling for epifauna richness}

# Check VIF for fixed effects
vif_model <- lm(Epifauna_Richness ~ Treatment + Water_Level + Veg_Cover, 
                data = combined_invert_data)
vif(vif_model) # these are very low, so this is good

# Conway-Maxwell-Poisson model for underdispersion
epi_richness_model <- glmmTMB(Epifauna_Richness ~ Treatment + Water_Level + Veg_Cover + (1 | Site),
                             data = combined_invert_data,
                             family = compois(link = "log"))

# Model diagnostics
model_resids <- simulateResiduals(epi_richness_model, plot = TRUE) # not perfect but acceptable

# Additional diagnostic tests
testDispersion(model_resids) #non-significant p-value is good
testZeroInflation(model_resids) #non-significant p-value is good
testOutliers(model_resids) #outlier test is non-significant

# Diagnostic plots
plot(model_resids$scaledResiduals ~ combined_invert_data$Veg_Cover,
     xlab = "Vegetation Cover (%)", ylab = "Scaled Residuals")
plot(model_resids$scaledResiduals ~ as.factor(combined_invert_data$Treatment),
     xlab = "Runnel", ylab = "Scaled Residuals")
plot(model_resids$scaledResiduals ~ as.factor(combined_invert_data$Water_Level),
     xlab = "Water_Level", ylab = "Scaled Residuals")

# Model summary and checks
summary(epi_richness_model)
confint(epi_richness_model)
r.squaredGLMM(epi_richness_model)


```

```{r glmmTMB for infauna richness}

# Check VIF for fixed effects
vif_model <- lm(Infauna_Richness ~ Treatment + Water_Level + Veg_Cover, 
                data = combined_invert_data)
vif(vif_model) # Very low (around 1) so these are good

# Conway-Maxwell-Poisson model for underdispersion
inf_richness_model <- glmmTMB(Infauna_Richness ~ Treatment + Water_Level + Veg_Cover + (1 | Site),
                             data = combined_invert_data,
                             family = compois(link = "log"))

# Model diagnostics
model_resids <- simulateResiduals(inf_richness_model, plot = TRUE) #both plots look good

# Additional diagnostic tests
testDispersion(model_resids) #non-significant
testZeroInflation(model_resids) #non-significant
testOutliers(model_resids) #no significant outliers

# Diagnostic plots - modified slightly to handle an NA
model_data <- model.frame(inf_richness_model)
plot_data <- data.frame(
  residuals = model_resids$scaledResiduals,
  veg_cover = model_data$Veg_Cover,
  treatment = model_data$Treatment,
  Water_Level = model_data$Water_Level
)

# Plot residuals vs vegetation cover
plot(plot_data$residuals ~ plot_data$veg_cover,
     xlab = "Vegetation Cover (%)", 
     ylab = "Scaled Residuals")

# Plot residuals vs runnel
boxplot(plot_data$residuals ~ as.factor(plot_data$treatment),
        xlab = "Treatment", 
        ylab = "Scaled Residuals")

# Plot residuals vs Water_Level
boxplot(plot_data$residuals ~ as.factor(plot_data$Water_Level),
        xlab = "Water_Level", 
        ylab = "Scaled Residuals")

# Model summary and checks
summary(inf_richness_model)
confint(inf_richness_model)
r.squaredGLMM(inf_richness_model)

```

#Invertebrate Diversity
```{r lme4 modelling for epifauna diversity}

epi_diversity_model <- lme4::lmer(Epifauna_Shannon ~ 
                                    Treatment + 
                                    Water_Level + 
                                    Veg_Cover + 
                                    (1 | Site),
                           data = combined_invert_data)

# Model diagnostics using DHARMa
model_resids <- simulateResiduals(epi_diversity_model, plot = TRUE) #Not looking good.

# Additional diagnostic tests
testDispersion(model_resids)
testZeroInflation(model_resids) # zero-inflated
testOutliers(model_resids)

#Moving on to using nonparametric bootstrapping

# Standard confidence intervals
cat("\nStandard Confidence Intervals:\n")
confint(epi_diversity_model)

# Bootstrapped confidence intervals
cat("\nBootstrapped Confidence Intervals:\n")
set.seed(123) # for reproducibility 
boot_ci <- confint(epi_diversity_model, method="boot", nsim=1000)
print(boot_ci)
# Note: Singular fit warnings in some bootstrap samples, but similar confidence intervals to standard method, and we don't see any change in the overlap with zero (i.e. did not previously overlap with zero and now does). 
# Suggests stable parameter estimates

# Variance components interpretation:
# .sig01: Random effect std dev (Site) confidence intervals includes 0 [0.067, 0.37] suggesting minimal site-level variation
# .sigma: Residual std dev. confidence interval [0.27, 0.37] represents unexplained variation within sites

# Compare standard vs bootstrapped intervals
cat("\nComparison of Confidence Intervals:\n")
ci_comparison <- data.frame(
  Parameter = rownames(boot_ci),
  Standard_Lower = confint(epi_diversity_model)[,1],
  Standard_Upper = confint(epi_diversity_model)[,2],
  Boot_Lower = boot_ci[,1],
  Boot_Upper = boot_ci[,2]
)
print(ci_comparison) #As noted before, the confidence intervals are similar which is good


# Model summary
summary(epi_diversity_model)

# Get p-values for fixed effects
summary(as_lmerModLmerTest(epi_diversity_model))  # converts to lmerTest object

# R-squared
r.squaredGLMM(epi_diversity_model)

# Effect plots of predicted relationships
epi_effects <- ggpredict(epi_diversity_model, terms = c("Water_Level", "Veg_Cover"))
plot(epi_effects)

```

```{r lme4 modeling for infauna diversity}

inf_diversity_model <- lme4::lmer(Infauna_Shannon ~ 
                                   Treatment + 
                                   Water_Level + 
                                   Veg_Cover + 
                                   (1 | Site),
                                 data = combined_invert_data)

# Model diagnostics using DHARMa
model_resids <- simulateResiduals(inf_diversity_model, plot = TRUE)

# Additional diagnostic tests
testDispersion(model_resids)
testZeroInflation(model_resids) #zero-inflated
testOutliers(model_resids)

# Moving on to bootstrapping.

confint(inf_diversity_model)

# Bootstrapped confidence intervals
set.seed(123) # for reproducibility 
boot_ci <- confint(inf_diversity_model, method="boot", nsim=1000)
print(boot_ci)
# Note: Singular fit warnings in some bootstrap samples, but similar confidence intervals to standard method 
# Suggests stable parameter estimates

# Variance components interpretation: see above

# Compare standard vs bootstrapped intervals
ci_comparison <- data.frame(
  Parameter = rownames(boot_ci),
  Standard_Lower = confint(inf_diversity_model)[,1],
  Standard_Upper = confint(inf_diversity_model)[,2],
  Boot_Lower = boot_ci[,1],
  Boot_Upper = boot_ci[,2]
)
print(ci_comparison) # good

# Model summary
summary(inf_diversity_model)

# Get p-values for fixed effects
summary(as_lmerModLmerTest(inf_diversity_model))  # converts to lmerTest object

# R-squared
r.squaredGLMM(inf_diversity_model)

# Effect plots of predicted relationships
inf_effects <- ggpredict(inf_diversity_model, terms = c("Water_Level", "Veg_Cover"))
plot(inf_effects)

```



# Fish

## Q1: Are there differences in fish abundance/size between runnel and non-runnel sites?
```{r runnel vs non-runnel fish abundance - platform only}

# Comparing runnel vs non-runnel sites using traps on the marsh platform only

# Mixed effects model for fish abundance
runnel_vs_control_abundance <- glmmTMB(Total_Number_Fish_by_Trap ~ 
                                         Treatment + 
                                         Tide +
                                         (1 | Site),
                                       data = fish_field_data_no_traps_in_runnels,
                                       family = nbinom2(link = "log"))

# Model diagnostics
model_resids <- simulateResiduals(runnel_vs_control_abundance, plot = TRUE)
# Additional diagnostic tests
testDispersion(model_resids) #non-significant p-value is good
testZeroInflation(model_resids) #non-significant p-value is good
testOutliers(model_resids) #outlier test is non-significant, 

# Model summary
summary(runnel_vs_control_abundance)
confint(runnel_vs_control_abundance)
r.squaredGLMM(runnel_vs_control_abundance)

# Create a prediction grid for abundance
abundance_preds <- ggpredict(runnel_vs_control_abundance, terms = c("Treatment", "Tide"))
abundance_preds <- as.data.frame(abundance_preds)
abundance_preds$Treatment <- factor(abundance_preds$x, levels = levels(fish_field_data_no_traps_in_runnels$Treatment))
abundance_preds$Tide <- factor(abundance_preds$group, levels = levels(fish_field_data_no_traps_in_runnels$Tide))


```

```{r runnel vs non-runnel fish weight - platform only}

# Comparing runnel vs non-runnel sites using traps on the marsh platform only

# Linear model for fish weight (no random effect due to super low variance)
runnel_vs_control_weight <- lm(log(Fish_Weight) ~ 
                                        Treatment + 
                                         Tide,
                                       data = fish_field_size_data_no_traps_in_runnels)

# Model diagnostics for linear model
par(mfrow = c(2, 2))
plot(runnel_vs_control_weight)
par(mfrow = c(1, 1))

# Model summary
summary(runnel_vs_control_weight)
r.squaredGLMM(runnel_vs_control_weight)

# Create prediction grid
pred_grid <- expand.grid(
  Treatment = c("Control", "Runnel"),
  Tide = c("H", "L")
)

# Get predictions on transformed scale
pred <- predict(runnel_vs_control_weight, newdata = pred_grid, re.form = ~0, 
                                      se.fit = TRUE)

# Store predictions and SE on transformed scale
pred_grid$pred_transformed <- pred$fit
pred_grid$se_transformed   <- pred$se.fit

# Back-transform to original scale
pred_grid$pred_weight <- exp(pred_grid$pred_transformed)

# Back-transform SE using delta method
pred_grid$se_weight <- pred_grid$se_transformed * exp(pred_grid$pred_transformed)

# Calculate means by treatment group
weight_effects_summary <- pred_grid %>%
  group_by(Treatment, Tide) %>%
  summarise(
    mean_weight = mean(pred_weight, na.rm = TRUE),
    se_weight   = mean(se_weight, na.rm = TRUE),  
    .groups = 'drop'
  )

```

```{r runnel vs non-runnel fish length - platform only}

# Linear model for fish length for platform traps only (no random effect due to super low variance)
runnel_vs_control_length <- lm(log(Fish_Length) ~ 
                                        Treatment + 
                                         Tide,
                                       data = fish_field_size_data_no_traps_in_runnels)

# Model diagnostics for linear model
par(mfrow = c(2, 2))
plot(runnel_vs_control_length)
par(mfrow = c(1, 1))

# Model summary
summary(runnel_vs_control_length)
r.squaredGLMM(runnel_vs_control_length)


# Get model predictions for length on the prediction grid
pred <- predict(runnel_vs_control_length, newdata = pred_grid, re.form = ~0, se.fit = TRUE)

pred_grid$pred_length <- pred$fit
pred_grid$se_length   <- pred$se.fit   

# Summarize for plotting
length_effects_summary <- pred_grid %>%
  group_by(Treatment, Tide) %>%
  summarise(
    mean_length = mean(pred_length, na.rm = TRUE),
    se_length   = mean(se_length, na.rm = TRUE),  
    .groups = 'drop'
  )

length_effects_summary$Treatment <- factor(length_effects_summary$Treatment, levels = c("Control", "Runnel"))
length_effects_summary$Tide <- factor(length_effects_summary$Tide, levels = c("H", "L"))

```

## Q2: Do fish use runnels? (within runnel-sites only)
```{r within-runnel sites, abundance comparison}

# Comparing fish caught inside runnels vs. on platforms, at runnel sites only

# Filter for runnel sites only
runnel_sites_data <- fish_field_data %>% filter(Treatment == "Runnel")

# Summary of trap locations and types for runnel sites
table(runnel_sites_data$Trap_Location, runnel_sites_data$Trap_Type)

# Mixed effects model for fish abundance within runnel sites
within_runnel_abundance <- glmmTMB(Total_Number_Fish_by_Trap ~ 
                                     Trap_Location * Tide +  
                                     (1 | Site),
                                   data = runnel_sites_data,
                                   family = nbinom2(link = "log"))

# Model diagnostics
model_resids <- simulateResiduals(within_runnel_abundance, plot = TRUE)
# Additional diagnostic tests
testDispersion(model_resids) #non-significant p-value is good
testZeroInflation(model_resids) #non-significant p-value is good
testOutliers(model_resids) #outlier test is non-significant

# Model summary
summary(within_runnel_abundance)
confint(within_runnel_abundance)
r.squaredGLMM(within_runnel_abundance)

# Create a prediction grid for abundance
abundance_preds <- ggpredict(within_runnel_abundance, terms = c("Trap_Location", "Tide"))
abundance_preds <- as.data.frame(abundance_preds)
abundance_preds$Trap_Location <- factor(abundance_preds$x, levels = levels(runnel_sites_data$Trap_Location))
abundance_preds$Tide <- factor(abundance_preds$group)

##Interpretation: Fish are using the runnels overall, we see more fish in the runnels. Interaction effect is not significant, meaning fish are using the runnels to get on and off the platform. Slightly higher site variation here, but that might be expected because we have fewer sites in this analysis.

```

```{r within-runnel sites, fish weight analysis}

# Comparing fish caught inside runnels vs. on platforms, at runnel sites only

# Filter for runnel sites only
runnel_fish_data <- fish_field_size_data %>% filter(Treatment == "Runnel")

# Linear model for fish weight (no random effect due to super low variance)
within_runnel_weight <- lm(I(Fish_Weight^(1/3)) ~ 
                             Trap_Location + 
                             Tide,
                           data = runnel_fish_data)

# Model diagnostics for linear model
par(mfrow = c(2, 2))
plot(within_runnel_weight)
par(mfrow = c(1, 1))

# Model summary
summary(within_runnel_weight)
cat("\nR-squared:", round(summary(within_runnel_weight)$r.squared, 3), "\n")

# Create a prediction grid for weight
runnel_fish_data$Tide <- as.factor(runnel_fish_data$Tide)
print(levels(runnel_fish_data$Trap_Location))
print(levels(runnel_fish_data$Tide))

# Creating the prediction grid
within_pred_grid <- expand.grid(
  Trap_Location = levels(runnel_fish_data$Trap_Location),
  Tide = levels(runnel_fish_data$Tide)
)

# Get predictions on transformed scale, including SE
pred <- predict(within_runnel_weight, newdata = within_pred_grid, se.fit = TRUE, re.form = ~0)

# Store predictions and model-based SE
within_pred_grid$pred_transformed <- pred$fit
within_pred_grid$se_transformed   <- pred$se.fit

# Back-transform predictions to original scale
within_pred_grid$pred_weight <- within_pred_grid$pred_transformed^3

# Back-transform SE using delta method
within_pred_grid$se_weight <- within_pred_grid$se_transformed * (3 * within_pred_grid$pred_transformed^2)

# Summarize by group
within_weight_effects_summary <- within_pred_grid %>%
  group_by(Trap_Location, Tide) %>%
  summarise(
    mean_weight = mean(pred_weight, na.rm = TRUE),
    se_weight   = mean(se_weight, na.rm = TRUE),  
    .groups = 'drop'
  )


```

```{r within-runnel sites, fish length analysis}

# Comparing fish caught inside runnels vs. on platforms, at runnel sites only

# Linear model for fish length (no random effect of site due to super low variance)
within_runnel_length <- lm(Fish_Length ~ 
                             Trap_Location + 
                             Tide,
                           data = runnel_fish_data)


# Model diagnostics for linear model
par(mfrow = c(2, 2))
plot(within_runnel_length)
par(mfrow = c(1, 1))

# Model summary
summary(within_runnel_length)
cat("\nR-squared:", round(summary(within_runnel_length)$r.squared, 3), "\n")


# Create prediction grid for length
within_length_pred_grid <- expand.grid(
  Trap_Location = levels(runnel_fish_data$Trap_Location),
  Tide = levels(runnel_fish_data$Tide)
)

# Get predictions and model-based SE
pred <- predict(within_runnel_length, newdata = within_length_pred_grid, 
                se.fit = TRUE, re.form = ~0)

# Store predictions and SE
within_length_pred_grid$pred_length <- pred$fit
within_length_pred_grid$se_length   <- pred$se.fit

# Summarize by group for plotting
within_length_effects_summary <- within_length_pred_grid %>%
  group_by(Trap_Location, Tide) %>%
  summarise(
    mean_length = mean(pred_length, na.rm = TRUE),
    se_length   = mean(se_length, na.rm = TRUE),  # use model-based SE
    .groups = 'drop'
  )

```

## Q3: Are fish coming onto the platform with empty stomachs?
```{r incoming vs receding tide stomach comparison}

# Comparing incoming (L) vs receding (H) tide fish to confirm "empty stomach"

# Filter for trap location to excludes traps in the runnels 
tide_comparison_data <- combined_fish_data %>%
  filter(Trap_Location != "Runnel")

# --- Model for Percent Digested ---

# Response 'Digested' is a percentage --> Beta regression
# We have such low variance in the random effect that is causes singularity issues. We're going to drop the random effects here and use `betareg`

# Convert 'Digested' to a proportion and transform to handle 0s and 1s
n <- nrow(tide_comparison_data)
tide_comparison_data <- tide_comparison_data %>%
  mutate(Digested_prop = Digested / 100,
         Digested_prop_transformed = (Digested_prop * (n - 1) + 0.5) / n)

# beta regression model
digested_model_beta <- betareg(Digested_prop_transformed ~ Tide, data = tide_comparison_data)

# Model diagnostics for beta regression; can't use DHARMa here
# 1. Standard residual plots
par(mfrow = c(2, 2))
plot(digested_model_beta, which = 1:4)

# Pearson residuals vs fitted values
fitted_vals <- fitted(digested_model_beta)
pearson_resid <- residuals(digested_model_beta, type = "pearson")

plot(fitted_vals, pearson_resid,
     xlab = "Fitted values",
     ylab = "Pearson residuals",
     main = "Pearson Residuals vs Fitted Values")
abline(h = 0, lty = 2)


par(mfrow = c(1, 1)) # reset plot layout

# Pseudo R-squared
pseudo_r2 <- 1 - (digested_model_beta$loglik / digested_model_beta$null.loglik)
cat("Pseudo R-squared:", round(pseudo_r2, 3), "\n") # low here but not surprising

# Comparison with null model
null_model <- betareg(Digested_prop_transformed ~ 1, data = tide_comparison_data)

# Likelihood ratio test
lr_stat <- 2 * (digested_model_beta$loglik - null_model$loglik)
df_diff <- length(coef(digested_model_beta)) - length(coef(null_model))
p_value <- 1 - pchisq(lr_stat, df_diff)

cat("Test statistic:", round(lr_stat, 3), "\n")
cat("Degrees of freedom:", df_diff, "\n")
cat("P-value:", round(p_value, 4), "\n")

# Model summary and confidence intervals
summary(digested_model_beta)
confint(digested_model_beta)

##Interpretation: The % digested as a beta regression showed (a) very little difference between sites and (b) a strong effect of tides. The "empty stomach" can be confirmed here; fish caught on the incoming tide (L) have a significantly higher proportion of digested material (coefficient of 1.1508) compared to fish caught on the receding tide 

```

## Q4: What do fish consume on the platform; is it different between runnelled and un-runnelled sites?

```{r receding tide NMDS analysis; full data}

# NMDS analysis of fish diet composition on receding tide

# Filter for receding tide data only
fish_diet_receding <- combined_fish_data %>% 
  filter(Tide == "H")

# Remove rows with 100% digested content (no diet data)
fish_diet_receding_no_100 <- fish_diet_receding %>%
  filter(Digested != "100")

# Remove any columns with all zeroes
fish_diet_receding_no_100 <- fish_diet_receding_no_100[, colSums(fish_diet_receding_no_100 != 0) > 0]

# Separate environmental data and species data
fish_diet_spp <- fish_diet_receding_no_100 %>% select(c(Algae:Beetles, Diptera:Total_Crab))

# Check for zero-variance columns and remove them
zero_var_cols <- sapply(fish_diet_spp, function(x) var(x, na.rm = TRUE) == 0)
if(any(zero_var_cols)) {
  cat("Removing zero-variance columns:\n")
  print(names(fish_diet_spp)[zero_var_cols])
  fish_diet_spp <- fish_diet_spp[, !zero_var_cols]
}

# Also check for very low variance columns
low_var_cols <- sapply(fish_diet_spp, function(x) {
  var_val <- var(x, na.rm = TRUE)
  return(var_val < 0.001)  # Very low variance threshold
})
if(any(low_var_cols)) {
  cat("Low-variance columns found:\n")
  print(names(fish_diet_spp)[low_var_cols])
}

# Environmental dataframe
fish_diet_env <- fish_diet_receding_no_100 %>% 
  select(c(Site:Tide, Fish_ID, Trap_Type:Treatment)) %>% 
  mutate(Site_ID_Tide_Version = paste(Site, ID, Tide, Fish_ID, sep = "")) %>% 
  select(-Site, -ID, -Tide, -Fish_ID) 

# Diagnostic information before NMDS (initial analysis)
cat("Data structure before outlier removal:\n")
cat("Number of samples:", nrow(fish_diet_spp), "\n")
cat("Number of species:", ncol(fish_diet_spp), "\n")
cat("Total abundance:", sum(fish_diet_spp, na.rm = TRUE), "\n")
cat("Number of zero entries:", sum(fish_diet_spp == 0, na.rm = TRUE), "\n")
cat("Proportion of zeros:", round(sum(fish_diet_spp == 0, na.rm = TRUE) / length(as.matrix(fish_diet_spp)), 3), "\n")

if(nrow(fish_diet_spp) < 3) {
  cat("WARNING: Too few samples for NMDS analysis\n")
} else if(ncol(fish_diet_spp) < 2) {
  cat("WARNING: Too few species for NMDS analysis\n")
} else {
  # Performing nmds with environmental (treatment) and species vectors
  fish_diet_mds <- metaMDS(fish_diet_spp, distance = "bray", autotransform = FALSE)
  
  # Check stress value
  cat("NMDS stress value:", round(fish_diet_mds$stress, 4), "\n")
}

# Only run envfit if NMDS was successful and we have species data
if(!inherits(fish_diet_mds, "try-error") && ncol(fish_diet_spp) > 0) {
  fish_diet_sppfit <- envfit(fish_diet_mds, fish_diet_spp, permutations = 999) # Fit species vectors
} else {
  cat("NMDS failed or no species data available. Check your data.\n")
}

fish_diet_mds  # Check stress value

# Saving initial NMDS results
fish_diet_site_scrs <- as.data.frame(scores(fish_diet_mds, display = "sites"))
fish_diet_site_scrs <- cbind(fish_diet_site_scrs, Treatment = fish_diet_env$Treatment)

## Checking for outliers

# Outliers identified as samples with NMDS1 scores above the 99th percentile in the initial NMDS ordination of fish diet composition.
outlier_indices <- which(fish_diet_site_scrs$NMDS1 > quantile(fish_diet_site_scrs$NMDS1, 0.99))
outlier_samples <- rownames(fish_diet_site_scrs)[outlier_indices]
print(outlier_samples)  # "LBSL3H36" "LBSL3H23"

# Remove rows with outliers
fish_diet_receding_no_100_no_outliers <- fish_diet_receding_no_100 %>%
  .[!(rownames(.) %in% outlier_samples), ]

fish_diet_receding_no_100_no_outliers <- fish_diet_receding_no_100_no_outliers[, colSums(fish_diet_receding_no_100 != 0) > 0]

## Re-running NMDS with outliers removed

# Separate environmental data and species data
fish_diet_spp <- fish_diet_receding_no_100_no_outliers %>% select(c(Algae:Beetles,Diptera:Total_Crab))

# Check for zero-variance columns and remove them (after outlier removal)
zero_var_cols_no_outliers <- sapply(fish_diet_spp, function(x) var(x, na.rm = TRUE) == 0)
if(any(zero_var_cols_no_outliers)) {
  cat("Removing zero-variance columns (after outlier removal):\n")
  print(names(fish_diet_spp)[zero_var_cols_no_outliers])
  fish_diet_spp <- fish_diet_spp[, !zero_var_cols_no_outliers]
}

# Also check for very low variance columns
low_var_cols_no_outliers <- sapply(fish_diet_spp, function(x) {
  var_val <- var(x, na.rm = TRUE)
  return(var_val < 0.001)  # Very low variance threshold
})
if(any(low_var_cols_no_outliers)) {
  cat("Low-variance columns found (after outlier removal):\n")
  print(names(fish_diet_spp)[low_var_cols_no_outliers])
}

# Environmental dataframe
fish_diet_env <- fish_diet_receding_no_100_no_outliers %>% 
  select(c(Site:Tide, Fish_ID, Trap_Type:Treatment)) %>% 
  mutate(Site_ID_Tide_Version = paste(Site, ID, Tide, Fish_ID, sep = "")) %>% 
  select(-Site, -ID, -Tide, -Fish_ID) 

# Diagnostic information before NMDS (after outlier removal)
cat("Data structure after outlier removal:\n")
cat("Number of samples:", nrow(fish_diet_spp), "\n")
cat("Number of species:", ncol(fish_diet_spp), "\n")
cat("Total abundance:", sum(fish_diet_spp, na.rm = TRUE), "\n")
cat("Number of zero entries:", sum(fish_diet_spp == 0, na.rm = TRUE), "\n")
cat("Proportion of zeros:", round(sum(fish_diet_spp == 0, na.rm = TRUE) / length(as.matrix(fish_diet_spp)), 3), "\n")

# Check if we have enough data for meaningful NMDS
if(nrow(fish_diet_spp) < 3) {
  cat("WARNING: Too few samples for NMDS analysis\n")
} else if(ncol(fish_diet_spp) < 2) {
  cat("WARNING: Too few species for NMDS analysis\n")
} else {
  # Performing nmds with environmental (treatment) and species vectors
  fish_diet_mds <- metaMDS(fish_diet_spp, distance = "bray", autotransform = FALSE)
  
  # Check stress value
  cat("NMDS stress value:", round(fish_diet_mds$stress, 4), "\n")
}

# Only run envfit if NMDS was successful and we have species data
if(!inherits(fish_diet_mds, "try-error") && ncol(fish_diet_spp) > 0) {
  fish_diet_sppfit <- envfit(fish_diet_mds, fish_diet_spp, permutations = 999) # Fit species vectors
} else {
  cat("NMDS failed or no species data available. Check your data.\n")
}

fish_diet_mds  # Check stress value

# Saving initial NMDS results
fish_diet_site_scrs <- as.data.frame(scores(fish_diet_mds, display = "sites"))
fish_diet_site_scrs <- cbind(fish_diet_site_scrs, Treatment = fish_diet_env$Treatment)

# Species contributions
fish_diet_spp_scrs <- as.data.frame(scores(fish_diet_sppfit, display = "vectors"))
fish_diet_spp_scrs <- cbind(fish_diet_spp_scrs, Species = rownames(fish_diet_spp_scrs))
fish_diet_spp_scrs <- cbind(fish_diet_spp_scrs, pval = fish_diet_sppfit$vectors$pvals)
sig_fish_diet_spp_scrs <- subset(fish_diet_spp_scrs, pval<=0.05)

# PERMANOVA (adonis2) for overall fish diet composition difference,
permanova_result_fish_diet_trap_type <- adonis2(fish_diet_spp ~ Treatment, data = fish_diet_env, permutations = 999, strata = fish_diet_env$Trap_Type)
print(permanova_result_fish_diet_trap_type)

# Additional analysis to understand Trap_Type effect
# First, let's see the distribution of Trap_Type in our data
cat("\nTrap Type distribution in fish diet analysis:\n")
print(table(fish_diet_env$Trap_Type))

cat("\nTrap Type by Treatment:\n")
print(table(fish_diet_env$Trap_Type, fish_diet_env$Treatment))

# Run PERMANOVA without strata to compare
permanova_result_fish_diet_no_strata <- adonis2(fish_diet_spp ~ Treatment, data = fish_diet_env, permutations = 999)
cat("\nPERMANOVA without strata (for comparison):\n")
print(permanova_result_fish_diet_no_strata)

# Calculate effect size (R²) for both models
r2_with_strata <- permanova_result_fish_diet_trap_type$R2[1]
r2_without_strata <- permanova_result_fish_diet_no_strata$R2[1]

cat("\nEffect size comparison:\n")
cat("R² with Trap_Type strata:", round(r2_with_strata, 4), "\n")
cat("R² without strata:", round(r2_without_strata, 4), "\n")
cat("Difference in R²:", round(r2_without_strata - r2_with_strata, 4), "\n")

# Test if Trap_Type itself has a significant effect on diet composition
permanova_trap_type_effect <- adonis2(fish_diet_spp ~ Trap_Type, data = fish_diet_env, permutations = 999)
cat("\nPERMANOVA testing Trap_Type effect on diet composition:\n")
print(permanova_trap_type_effect)

# Calculate average dissimilarity within and between Trap_Type groups
trap_type_groups <- fish_diet_env$Trap_Type
trap_type_dissimilarity <- vegdist(fish_diet_spp, method = "bray")

# Calculate average dissimilarity within each Trap_Type
within_dissimilarity <- sapply(unique(trap_type_groups), function(trap_type) {
  group_indices <- which(trap_type_groups == trap_type)
  if(length(group_indices) > 1) {
    group_dist <- as.matrix(trap_type_dissimilarity)[group_indices, group_indices]
    mean(group_dist[upper.tri(group_dist)], na.rm = TRUE)
  } else {
    NA
  }
})

cat("\nAverage dissimilarity within Trap_Type groups:\n")
for(i in seq_along(within_dissimilarity)) {
  if(!is.na(within_dissimilarity[i])) {
    cat(names(within_dissimilarity)[i], ":", round(within_dissimilarity[i], 4), "\n")
  }
}

# Calculate overall average dissimilarity
overall_dissimilarity <- mean(as.matrix(trap_type_dissimilarity)[upper.tri(trap_type_dissimilarity)], na.rm = TRUE)
cat("Overall average dissimilarity:", round(overall_dissimilarity, 4), "\n")

# Calculate coefficient of variation for dissimilarity within groups
cv_within <- sapply(unique(trap_type_groups), function(trap_type) {
  group_indices <- which(trap_type_groups == trap_type)
  if(length(group_indices) > 1) {
    group_dist <- as.matrix(trap_type_dissimilarity)[group_indices, group_indices]
    group_vals <- group_dist[upper.tri(group_dist)]
    sd(group_vals, na.rm = TRUE) / mean(group_vals, na.rm = TRUE)
  } else {
    NA
  }
})

cat("\nCoefficient of variation for dissimilarity within Trap_Type groups:\n")
for(i in seq_along(cv_within)) {
  if(!is.na(cv_within[i])) {
    cat(names(cv_within)[i], ":", round(cv_within[i], 4), "\n")
  }
}

# SIMPER analysis for fish diet
mat_bar_fish_diet_spp <- as.matrix(fish_diet_spp)
simp_env_fish_diet <- simper(mat_bar_fish_diet_spp, fish_diet_env$Treatment)
summary(simp_env_fish_diet, ordered = TRUE, digits = 3)

# Additional diagnostic: NMDS plot colored by Trap_Type to visualize the effect
fish_diet_site_scrs_with_trap_type <- cbind(fish_diet_site_scrs, Trap_Type = fish_diet_env$Trap_Type)
```

## SQ1: Does trap type affect fish dynamics?

```{r fish abundance between trap types - control sites only}

# Comparing fish caught in Breder versus bottle traps at control sites only, where all traps were placed on the platform

# Filter for control sites only
control_sites_data <- fish_field_data %>% filter(Treatment == "Control")

# Summary of trap locations and types for runnel sites
table(control_sites_data$Trap_Type)

# Mixed effects model for fish abundance within control sites
within_control_abundance <- glmmTMB(Total_Number_Fish_by_Trap ~ 
                                     Trap_Type +
                                     (1 | Site),
                                   data = control_sites_data,
                                   family = nbinom2(link = "log"))

# Model diagnostics
model_resids <- simulateResiduals(within_control_abundance, plot = TRUE)
# Additional diagnostic tests
testDispersion(model_resids) #non-significant p-value is good
testZeroInflation(model_resids) #non-significant p-value is good
testOutliers(model_resids) #outlier test is non-significant, indicating no significant outliers 

# Model summary
summary(within_control_abundance)
confint(within_control_abundance)
r.squaredGLMM(within_control_abundance)

# Create a prediction grid for abundance with Trap_Type only
abundance_preds <- ggpredict(within_control_abundance, terms = c("Trap_Type"))
abundance_preds <- as.data.frame(abundance_preds)
abundance_preds$Trap_Type <- factor(abundance_preds$x, levels = levels(control_sites_data$Trap_Type))

```

```{r fish weight between trap types - control sites only}

# Comparing fish caught in Breder versus bottle traps

control_fish_data <- fish_field_size_data %>% filter(Treatment == "Control")

# Model with no random effect due to super low variance
within_control_weight <- lm(log(Fish_Weight) ~ 
                                         Trap_Type,
                                       data = control_fish_data)

# Model diagnostics for linear model
par(mfrow = c(2, 2))
plot(within_control_weight)
par(mfrow = c(1, 1))

# Model summary
summary(within_control_weight)
r.squaredGLMM(within_control_weight)

# Create prediction grid
pred_grid <- expand.grid(
  Trap_Type = c("Breder", "Bottle")
)

# Get predictions on transformed scale, including model-based SE
pred <- predict(within_control_weight, newdata = pred_grid, re.form = ~0, se.fit = TRUE)

# Store predictions and SE
pred_grid$pred_transformed <- pred$fit
pred_grid$se_transformed   <- pred$se.fit

# Back-transform to original scale (log -> original)
pred_grid$pred_weight <- exp(pred_grid$pred_transformed)

# Back-transform SE using delta method
pred_grid$se_weight <- pred_grid$se_transformed * exp(pred_grid$pred_transformed)

# Summarize by Trap_Type for plotting
weight_effects_summary <- pred_grid %>%
  group_by(Trap_Type) %>%
  summarise(
    mean_weight = mean(pred_weight, na.rm = TRUE),
    se_weight   = mean(se_weight, na.rm = TRUE),  # use model-based SE
    .groups = 'drop'
  )

##Interpretation: trap type is not significant for weight.

```

```{r fish length between trap types - control sites only}

# Comparing fish caught in Breder versus bottle traps

# Model with no random effect due to super low variance
within_control_length <- lm(log(Fish_Length) ~ 
                             Trap_Type,
                           data = control_fish_data)


# Model diagnostics for linear model
par(mfrow = c(2, 2))
plot(within_control_length)
par(mfrow = c(1, 1))

# Model summary
summary(within_control_length)
r.squaredGLMM(within_control_length)

# Create prediction grid
pred_grid <- expand.grid(
  Trap_Type = levels(control_fish_data$Trap_Type)
)

# Get predictions and model-based SE
pred <- predict(within_control_length, newdata = pred_grid, re.form = ~0, se.fit = TRUE)

# Store predictions and SE
pred_grid$pred_length <- pred$fit
pred_grid$se_length   <- pred$se.fit

# Summarize by Trap_Type for plotting
length_effects_summary <- pred_grid %>%
  group_by(Trap_Type) %>%
  summarise(
    mean_length = mean(pred_length, na.rm = TRUE),
    se_length   = mean(se_length, na.rm = TRUE),  # use model-based SE
    .groups = 'drop'
  )

```

